---
title: "R Notebook"
output: html_notebook
---
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

```{r}
library(ISLR2)
Gitters <- na.omit(Hitters)
n <- nrow(Gitters)
set.seed(13) 
ntest <- trunc(n / 3)
testid <- sample(1:n, ntest)
```

```{r}
#NB: Note the special tilde-character
lfit <- lm(Salary ~ ., data = Gitters[-testid, ])
```

```{r}
lpred <- predict(lfit, Gitters[testid, ])
with(Gitters[testid, ], mean(abs(lpred - Salary)))
```

```{r}
x <- scale(model.matrix(Salary ~ . - 1, data = Gitters))
y <- Gitters$Salary
```

```{r}
library(glmnet)
cvfit <- cv.glmnet(x[-testid, ], y[-testid],
type.measure = "mae")
cpred <- predict(cvfit, x[testid, ], s = "lambda.min")
mean(abs(y[testid] - cpred))
```

```{r}
library(keras)
modnn <- keras_model_sequential() %>%
  layer_dense(units = 50, activation = "relu",
  input_shape = ncol(x)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units = 1)
```

```{r}
x <- scale(model.matrix(Salary ~ . - 1, data = Gitters))
```

```{r}
x <- model.matrix(Salary ~ . - 1, data = Gitters) %>% scale()
```

```{r}
modnn %>% compile(loss = "mse",
  optimizer = optimizer_rmsprop(),
  metrics = list("mean_absolute_error")
)
```

```{r}
history <- modnn %>% fit(
  x[-testid, ], y[-testid], epochs = 1500, batch_size = 32,
  validation_data = list(x[testid, ], y[testid])
)
```
```{r}
npred <- predict(modnn, x[testid, ])
mean(abs(y[testid] - npred))
```

```{r}
mnist <- dataset_mnist()
x_train <- mnist$train$x
g_train <- mnist$train$y
x_test <- mnist$test$x
g_test <- mnist$test$y
dim(x_train)
dim(x_test)
```

```{r}
x_train <- array_reshape(x_train, c(nrow(x_train), 784))
x_test <- array_reshape(x_test, c(nrow(x_test), 784))
y_train <- to_categorical(g_train, 10)
y_test <- to_categorical(g_test, 10)
```

```{r}
x_train <- x_train / 255
x_test <- x_test / 255
```

```{r}
modelnn <- keras_model_sequential()
modelnn %>% layer_dense(units = 256, activation = "relu", input_shape = c(784)) %>%
  layer_dropout(rate = 0.4) %>%
  layer_dense(units = 128, activation = "relu") %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = "softmax")
```

```{r}
summary(modelnn)
```

```{r}
modelnn %>% compile(loss = "categorical_crossentropy",
  optimizer = optimizer_rmsprop(), metrics = c("accuracy")
)
```

```{r}
system.time(
  history <- modelnn %>%
  fit(x_train, y_train, epochs = 30, batch_size = 128, validation_split = 0.2)
)
plot(history, smooth = FALSE)
```

```{r}
accuracy <- function(pred, truth)
  mean(drop(as.numeric(pred)) == drop(truth))
modelnn %>% predict(x_test) %>% k_argmax() %>% accuracy(g_test)
```

```{r}
modellr <- keras_model_sequential() %>%
  layer_dense(input_shape = 784, units = 10, activation = "softmax")
summary(modellr)
```

```{r}
modellr %>% compile(loss = "categorical_crossentropy", optimizer = optimizer_rmsprop(), metrics = c("accuracy"))
modellr %>% fit(x_train, y_train, epochs = 30, batch_size = 128, validation_split = 0.2)
modellr %>% predict(x_test) %>% k_argmax() %>% accuracy(g_test)
```

```{r}
cifar100 <- dataset_cifar100()
names(cifar100)
x_train <- cifar100$train$x
g_train <- cifar100$train$y
x_test <- cifar100$test$x
g_test <- cifar100$test$y
dim(x_train)
range(x_train[1,,, 1])
```

```{r}
x_train <- x_train / 255
x_test <- x_test / 255
y_train <- to_categorical(g_train, 100)
dim(y_train)
```

```{r}
library(jpeg)
par(mar = c(0, 0, 0, 0), mfrow = c(5, 5))
index <- sample(seq(50000), 25)
for (i in index) plot(as.raster(x_train[i,,, ]))
```

```{r}
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), padding = "same", activation = "relu", input_shape = c(32, 32, 3)) %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_conv_2d(filters = 256, kernel_size = c(3, 3), padding = "same", activation = "relu") %>%
  layer_max_pooling_2d(pool_size = c(2, 2)) %>%
  layer_flatten() %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 512, activation = "relu") %>%
  layer_dense(units = 100, activation = "softmax")
summary(model)
```

```{r}
model %>% compile(loss = "categorical_crossentropy", optimizer = optimizer_rmsprop(), metrics = c("accuracy"))
history <- model %>% fit(x_train, y_train, epochs = 30, batch_size = 128, validation_split = 0.2)
model %>% predict(x_test) %>% k_argmax() %>% accuracy(g_test)
```

```{r}
img_dir <- "/Users/esten/Downloads/book_images"
image_names <- list.files(img_dir)
num_images <- length(image_names)
x <- array(dim = c(num_images, 224, 224, 3))
for (i in 1:num_images) {
  img_path <- paste(img_dir, image_names[i], sep = "/")
  img <- image_load(img_path, target_size = c(224, 224))
  x[i,,, ] <- image_to_array(img)
}
x <- imagenet_preprocess_input(x)
```

```{r}
model <- application_resnet50(weights = "imagenet")
summary(model)
```

```{r}
pred6 <- model %>% predict(x) %>%
  imagenet_decode_predictions(top = 3)
names(pred6) <- image_names
print(pred6)
```

